---
title: 'Data types: Data science in a box '
author: Wesnahika
date: '2022-02-21'
slug: data-types-data-science-in-a-box
categories: []
tags:
  - data
  - data-science
  - r
---
# Data types

## Cat lovers 

A survey asked respondents their name and number of cats. The instructions said to enter the number of cats as a numerical value.

```{r}
library(tidyverse)
```

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/cat-lovers.csv'

cat_lovers <- read_csv(url)
```

### Oh why wont you work?!

```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats))
```
### mean

```{r}
?mean
```

### Oh why wont you still work??!!

```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats, na.rm = TRUE))
```

### look at the data 

```{r}
glimpse(cat_lovers)
```

### we might need to fix the data

```{r}
cat_lovers %>%
  mutate(number_of_cats = case_when(
    name == "Ginger Clark" ~ 2,
    name == "Doug Bass"    ~ 3,
    TRUE                   ~ as.numeric(number_of_cats)
    )) %>%
  summarise(mean_cats = mean(number_of_cats))
```

### you need to respect data types 

```{r}
cat_lovers %>%
  mutate(
    number_of_cats = case_when(
      name == "Ginger Clark" ~ "2",
      name == "Doug Bass"    ~ "3",
      TRUE                   ~ number_of_cats
      ),
    number_of_cats = as.numeric(number_of_cats)
    ) %>%
  summarise(mean_cats = mean(number_of_cats))
```

### Now everything is coming together

```{r}
cat_lovers <- cat_lovers %>%
  mutate(
    number_of_cats = case_when(
      name == "Ginger Clark" ~ "2",
      name == "Doug Bass"    ~ "3",
      TRUE                   ~ number_of_cats
      ),
    number_of_cats = as.numeric(number_of_cats)
    )
```

## Essentially...

If your data does not behave how you expect it to, type coercion upon reading in the data might be the reason.
Go in and investigate your data, apply the fix, save your data, live happily ever after.

# Data types in R

## Logical & Character 

### Logical 

Logical: - boolean values TRUE and FALSE

```{r}
typeof(TRUE)
```

### Character 

Character - character strings 

```{r}
typeof("hello")
```
## Double & integer

### Double

double - floating point numeric values (default numeric type)

```{r}
typeof(1.335)
```

```{r}
typeof(7)
```

### Integer

integer - integer numerical values (indicated with an L)

```{r}
typeof(7L)
```

```{r}
typeof(1:3)
```
## Concatenation 

Vectors can be constructed using the c() function.

```{r}
c(1,2,3)
```
```{r}
c("hello", "world!")
```

```{r}
c(c("hi", "hello"), c("bye", "jello"))
```

## Converting between types 

### With intention

```{r}
x <- 1:3
x
```

```{r}
typeof(x)
```

```{r}
y <- as.character(x)
y
```

```{r}
typeof(y)
```

```{r}
x <- c(TRUE, FALSE)
x
```

```{r}
typeof(x)
```
```{r}
y <- as.numeric(x)
y
```

```{r}
typeof(y)
```

### without intention

```{r}
c(1, "hello")
```

```{r}
c(FALSE, 3L)
```

```{r}
c(1.3, 3L)
```

```{r}
c(2L, "two")
```

## Explicit vs. implicit coercion

Explicit coercion is when you call a function like as.logical(), as.numeric(), as.integer(), as.double(), or as.character()

Implicit coercion happens when you use a vector in a specific context that expects a certain type of vector

## Special values 

NA: Not available
NaN: Not a number
Inf: Positive infinity
-Inf: Negative infinity

```{r}
pi/0
```

```{r}
0/0
```

```{r}
1/0 - 1/0
```

```{r}
1/0 + 1/0
```

### NAs are special 

```{r}
x <- c(1,2,3,4,NA)
mean(x)
```

```{r}
mean(x, na.rm = TRUE)
```

```{r}
summary(x)
```

### NAs are logical 

```{r}
typeof(NA)
```

### Mental model for NAs

Unlike NaN, NAs are genuinely unknown values
But that doesn't mean they can't function in a logical way
Let's think about why NAs are logical...

```{r}
TRUE | NA
```

```{r}
FALSE | NA
```

#### NA is unkonown, so it could be TRUE or FALSE 

```{r}
TRUE | TRUE
```

```{r}
FALSE | TRUE
```

```{r}
TRUE | FALSE 
```

```{r}
FALSE | FALSE
```

