---
title: 'Single dataframe operations '
author: ''
date: '2022-02-14'
slug: single-dataframe-operations
categories: []
tags: []
---
## Import file from git hub

### Call libraries 
```{r}
library(tidyverse)
```

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/hotels.csv'
hotels <- read_csv(url)
```

## Select

```
### Select a range of variables 
```{r}
hotels %>%
  select(hotel:arrival_date_month)
```

### Select variables with certain characteristics 

```{r}
hotels %>%
  select(starts_with("arrival"))
```

```{r}
hotels %>%
  select(ends_with("type"))
```
## Arrange

### Arrange in ascending and descending order 

```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(babies)
```

```{r}
hotels %>%
  select(adults, children, babies) %>%
  arrange(desc(babies))
```

## Slice

### Slice for certain row numbers 

```{r}
hotels %>%
  slice(1:5)
```

## Filter

### Filter to select a subset of rows 

```{r}
hotels %>%
  filter(hotel == "City Hotel")
```

### Filter for many conditions at once 

```{r}
hotels %>%
  filter( 
    adults == 0,
    children >= 1
    ) %>% 
  select(adults, babies, children)
```
### Filter for more complex conditions 

```{r}
hotels %>%
  filter( 
    adults == 0,     
    children >= 1 | babies >= 1     # | means or
    ) %>%
  select(adults, babies, children)
```
## Distinct 

### Distinct to filter for unique rows 

```{r}
hotels %>% 
  distinct(market_segment) %>%
  arrange(market_segment)
```
```{r}
hotels %>% 
  distinct(hotel, market_segment) %>%
  arrange(hotel, market_segment)
```

## Count 

### Count to create frequency tables 

```{r}
hotels %>%
  count(market_segment)
```

```{r}
hotels %>%
  count(market_segment, sort = TRUE)
```

### Count and arrange 
```{r}
hotels %>%
  count(market_segment) %>%
  arrange(n)
```

```{r}
hotels %>%
  count(market_segment) %>%
  arrange(desc(n))
```

### Count for multiple variables 

```{r}
hotels %>%
  count(hotel, market_segment)
```

### Order matters when you count 

```{r}
hotels %>%
  count(hotel, market_segment)
```

```{r}
hotels %>%
  count(market_segment, hotel)
```

## Muttate 

### Mutate to add a new variable 

```{r}
hotel_add_new <- hotels %>%
  mutate(little_ones = children + babies) %>%
  select(children, babies, little_ones) %>%
  arrange(desc(little_ones))

hotel_add_new
```

### Little ones in resort and city hotels 

```{r}
Hotel_Resort <- hotels %>%
  mutate(little_ones = children + babies) %>%
  filter(
    little_ones >= 1,
    hotel == "Resort Hotel"
    ) %>%
  select(hotel, little_ones)

Hotel_Resort
```

```{r}
Hotel_City <- hotels %>%
  mutate(little_ones = children + babies) %>%
  filter(
    little_ones >= 1,
    hotel == "City Hotel"
    )  %>%
  select(hotel, little_ones)

Hotel_City
```

## Summarize 

### summarize for summary stats 

summarise() changes the data frame entirely, it collapses rows down to a single summary statistic, and removes all columns that are irrelevant to the calculation.

```{r}
hotels %>%
  summarise(mean_adr = mean(adr))
```

### Multiple summary statististics 

summarise can be used for multiple summary statistics as well

```{r}
hotels %>%
  summarise(
    min_adr = min(adr),
    mean_adr = mean(adr),
    median_adr = median(adr),
    max_adr = max(adr)
    )
```

## Group by 

### Group by for group operations 

```{r}
hotels %>%
  group_by(hotel) %>%
  summarise(mean_adr = mean(adr))
```

### Calculating frequencies 

```{r}
hotels %>%
  group_by(hotel) %>%
  summarise(n = n())
```


