---
title: 'Professors Evaluations '
author: 'Wesnahika '
date: '2022-03-01'
slug: professors-evaluations
categories: []
tags:
  - GGPLOT2
  - data
  - data-science
  - r
---
## Preliminaries 

## Importing libraries and Data 

```{r}
library(moderndive)
library(tidyverse)
library(skimr)
library(gapminder)
```

##  Importing data 

```{r}
evals_ch5 <- evals %>%
  select(ID, score,bty_avg, age)
```

## Explore the Data 

```{r}
glimpse(evals_ch5)
```

## Sample the Data

```{r}
evals_ch5 %>%
  sample_n(size = 5)
```

## Mean and Median of BTY and SCore 
```{r}
evals_ch5 %>%
  summarize(mean_bty_avg = mean(bty_avg), mean_score = mean(score),
            median_bty_avg = median(bty_avg), median_score = median(score))
```

## Five number summary 

```{r}
evals_ch5 %>% select(score, bty_avg) %>% skim()
```

## Scatter plot of score vs BTY average
```{r}
plot(evals_ch5$score, evals_ch5$bty_avg)
```

## Correlation Between score and  BTY score 

```{r}
evals_ch5 %>% 
  get_correlation(formula = score ~ bty_avg)

evals_ch5 %>% 
  summarize(correlation = cor(score, bty_avg))
```

## Scatter Plot: Score vs BTY score 

```{r}
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", 
       y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

## Scatter Plot with Jitter 

```{r}
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

## Linear regression line on scatter plot

```{r}
ggplot(evals_ch5, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship between teaching and beauty scores") +  
  geom_smooth(method = "lm", se = FALSE)
```

## Linear regression 

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch5)
# Get regression table:
get_regression_table(score_model)
```

## Fitting the Regression model 

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch5)
# Get regression table:
get_regression_table(score_model)
```

## Regression points 

```{r}
regression_points <- get_regression_points(score_model)
regression_points
```

