---
title: 'Data classes: Data science in a box '
author: Wesnahika
date: '2022-02-23'
slug: data-classes-data-science-in-a-box
categories: []
tags:
  - data
  - data-science
  - dataframe
  - tidyverse
  - r
---
# Data classes 

We talked about types so far, next we'll introduce the concept of classes

Vectors are like Lego building blocks
We stick them together to build more complicated constructs, e.g. representations of data
The class attribute relates to the S3 class of an object which determines its behaviour
You don't need to worry about what S3 classes really mean, but you can read more about it here if you're curious
Examples: factors, dates, and data frames

## Factors 

R uses factors to handle categorical variables, variables that have a fixed and known set of possible values

```{r}
x <- factor(c("BS", "MS", "PhD", "MS"))
x
```

```{r}
typeof(x)
```

```{r}
class(x)
```

### More on factors 

```{r}
library(tidyverse)
glimpse(x)
```


```{r}
as.integer(x)
```

## Dates 

```{r}
y <- as.Date("2020-01-01")
y
```
```{r}
typeof(y)
```

```{r}
class(y)
```

### More on dates 

We can think of dates like an integer (the number of days since the origin, 1 Jan 1970) and an integer (the origin) glued together

```{r}
as.integer(y)
```
```{r}
as.integer(y)/365
```

## Data frames 

We can think of data frames like like vectors of equal length glued together

```{r}
df <- data.frame(x = 1:2, y = 3:4)
df
```

```{r}
typeof(df)
```

```{r}
class(df)
```
## Lists

```{r}
l <- list(
  x = 1:4,
  y = c("hi", "hello", "jello"),
  z = c(TRUE, FALSE)
)
l
```

## lists and data frames 

A data frame is a special list containing vectors of equal length
When we use the pull() function, we extract a vector from the data frame

```{r}
df
```

```{r}
df %>%
  pull(y)
```

# Working with factors 

## read the data using char strings 

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/cat-lovers.csv'

cat_lovers <- read_csv(url)
```

```{r}
library(tidyverse)
glimpse(cat_lovers)
```

## But coerce when plotting 

```{r}
ggplot(cat_lovers, mapping = aes( x = handedness)) +
  geom_bar()
```

## Forcats is used to manipulate factors 

```{r}
cat_lovers %>%
  mutate(handedness = fct_infreq(handedness)) %>%
  ggplot(mapping = aes(x = handedness)) + 
  geom_bar()
```

## points on factors 

Factors are useful when you have true categorical data and you want to override the ordering of character vectors to improve display.
They are also useful in modeling scenarios.
The forcats package provides a suite of useful tools that solve common problems with factors.

# Working with dates 

## How to make a date 

lubridate is the tidyverse-friendly package that makes dealing with dates a little easier.
It's not one of the core tidyverse packages, hence it's installed with install.packages("tidyverse) but it's not loaded with it, and needs to be explicitly loaded with library(lubridate).

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/hotels.csv' 
hotels <- read_csv(url)
```


```{r}
hotels %>%
  select(starts_with("arrival"))
```

### Step 1: Construct the class

```{r}
library(glue)

hotels %>%
  mutate(
    arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")
    ) %>% 
  relocate(arrival_date)
```

### Step 2: Count the bookings per date 

```{r}
hotels %>%
  mutate(arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")) %>%
  count(arrival_date)
## # A tibble: 793 x 2
```

### Step 3: Visualize bookings per date 

```{r}
hotels %>%
  mutate(arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")) %>%
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_line()
```

### looking a little harder... 

Why does the plot start with August when we know our data start in July? And why does 10 August come after 1 August?

### Step 1: Revise the construct dates 

```{r}
library(lubridate)
hotels %>%
  mutate(
    arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))
    ) %>% 
  relocate(arrival_date)
```

### Step 2: Count the bookings per date 

```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date)
```

### Step 3a: Visualize booking per date 

```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_line()
```

### Step 3b: Visualize the data with a smooth curve 

```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_smooth()
```

