---
title: 'Data IO '
author: Wesnahika
date: '2022-02-23'
slug: data-io
categories: []
tags:
  - data
  - data-science
  - r
---
# Importing data 

## Reading rectangular data into R

### readr

read_csv() - comma delimited files.
read_csv2() - semicolon separated files (common in countries where , is used as the decimal place).
read_tsv() - tab delimited files.
read_delim() - reads in files with any delimiter.
read_fwf() - fixed width files.
...

### readxl

read_excel() - read xls or xlsx files
...

## Reading data 
```{r}
library(tidyverse)
```


```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/nobel.csv'

nobel <- read_csv(url)
nobel
```

## Writing data 

```{r}
df <- tribble(
  ~x, ~y,
  1,  "a",
  2,  "b",
  3,  "c"
)

```

```{r}
read_csv("https://raw.githubusercontent.com/Wesnahika/tidyverse/main/df.csv")

```

# Variable names 

## data with bad names

```{r}
url <- "https://raw.githubusercontent.com/Wesnahika/tidyverse/main/edibnb-badnames.csv"
edibnb_badnames <- read_csv(url)
names(edibnb_badnames)
```

... but R doesn't allow spaces in variable names

### Option 1 - Define column names 

```{r}
edibnb_col_names <- read_csv(url,
                             col_names = c("id", "price", 
                                           "neighbourhood", "accommodates",
                                           "bathroom", "bedroom", 
                                           "bed", "review_scores_rating", 
                                           "n_reviews", "url"))
names(edibnb_col_names)
```

### Option 2 - Format text to snake_case

```{r}
edibnb_clean_names <- read_csv(url) %>%
  janitor::clean_names()
names(edibnb_clean_names)
```
 
# Variable types 

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/df-na.csv'

df_na <- read_csv(url)

df_na
```

### Option 1: Explicit NAs

```{r}
read_csv(url, 
         na = c("", "NA", ".", "9999", "Not applicable"))
```

### Option 2: Specify the column types 

```{r}
read_csv(url, col_types = list(col_double(), 
                                            col_character(), 
                                            col_character()))
```

# Case study: Favorite foods

## favorite foods 

```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/Wesnahika/tidyverse/main/favourite-food.csv")
fav_food
```

```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/Wesnahika/tidyverse/main/favourite-food.csv") %>%
  janitor::clean_names()

fav_food
```

## Handling NAs 

```{r}
fav_food <- read_csv('https://raw.githubusercontent.com/Wesnahika/tidyverse/main/favourite-food.csv',
                       na = c("N/A", "99999")) %>%
  janitor::clean_names()
fav_food
                       
```

## Make age numeric

```{r}
fav_food <- fav_food %>%
  mutate(
    age = if_else(age == "five", "5", age),
    age = as.numeric(age)
    )
glimpse(fav_food)
```

## What is the socio economic status? 

```{r}
fav_food %>%
  count(ses)
```

## Turn ses into a factor 

```{r}
fav_food <- fav_food %>%
  mutate(ses = fct_relevel(ses, "Low", "Middle", "High"))
fav_food %>%
  count(ses)
```

## Placing it together 

```{r}
fav_food <- read_csv('https://raw.githubusercontent.com/Wesnahika/tidyverse/main/favourite-food.csv') %>%
  janitor::clean_names() %>%
  mutate(
    age = if_else(age == "five", "5", age), 
    age = as.numeric(age),
    ses = fct_relevel(ses, "Low", "Middle", "High")
  )
```

## Out and back in 

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/fav-food-clean.csv'

fav_food_clean <- read_csv(url)
```

```{r}
fav_food_clean %>%
  count(ses)
```

# Other types of data 

googlesheets4: Google Sheets
haven: SPSS, Stata, and SAS files
DBI, along with a database specific backend (e.g. RMySQL, RSQLite, RPostgreSQL etc): allows you to run SQL queries against a database and return a data frame
jsonline: JSON
xml2: xml
rvest: web scraping
httr: web APIs
sparklyr: data loaded into spark