---
title: 'Working with multiple data frames '
author: ''
date: '2022-02-15'
slug: working-with-multiple-data-frames
categories: []
tags:
  - dataframe
---

# Data: women in science  '

## Call library 
```{r}
library(tidyverse)
```

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/professions.csv' 
professions = read_csv(url)
```

```{r}
url1 <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/works.csv'
works  = read_csv(url1)
```

```{r}
url2 <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/dates.csv'
dates = read_csv(url2)
```

## Inputs 
```{r}
professions
```
### Reminder 
```{r}
names(professions)
```
```{r}
names(dates)
```
```{r}
names(works)
```
```{r}
nrow(professions)
```

```{r}
nrow(dates)
```

```{r}
nrow(works)
```

## Joining data frames

### left_join()
```{r}
professions %>% 
  left_join(dates)
```

### right_join()

```{r}
professions %>%
  right_join(dates)
```

### full_join()

```{r}
dates %>%
  full_join(works)
```

### inner_join()

```{r}
dates %>%
  inner_join(works)
```

### Semi_join(works)

```{r}
dates %>%
  semi_join(works)

```

### anti_join()

```{r}
dates %>%
  anti_join(works)
```
### putting it together 

```{r}
professions %>%
  left_join(dates) %>%
  left_join(works)
```
# Case Study: Student records 

## Student records 
Have:
Enrolment: official university enrolment records
Survey: Student provided info missing students who never filled it out and including students who filled it out but dropped the class
Want: Survey info for all enrolled in class

## Just convert csv file

```{r}
url <- 'https://raw.githubusercontent.com/Wesnahika/tidyverse/main/enrolment.csv'
enrolment = read_csv(url)
```

```{r}
url1 <- 'https://github.com/Wesnahika/tidyverse/blob/main/survey.csv'
survey = read_csv(url1)
```

### In class
```{r}
enrolment %>%
  left_join(survey, by = "id")
```

### Survey missing 

```{r}
enrolment %>%
  anti_join(survey, by = "id")
```

### Dropped

```{r}
anti_join(survey, enrolment, by = "id")
```

